The increasing cost of happiness

### 2021-06-02 src/draft-SSM1.Rmd
Replace wealth with income (closes issue #6)

A fundamental question for society is how much happiness does a dollar buy? The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing wealth. While the location of this change point reportedly ranges between USD\$60K to \$95K, there has been no investigation as to whether this has increased or decreased over time. We tested the temporal relationship between income and affective wellbeing (happiness), and income and cognitive wellbeing (life satisfaction), using household economic data from Australia between 2001-2019. We discovered the change point between happiness and income has increased over those 19 years faster than inflation (i.e., cost of living). This suggests that inequalities in income may be driving increasing inequities in happiness between the rich and the poor, with implications for health and recent government policy-goals to monitor and improve wellbeing.

@@

Australia has low levels of income disparity relative to many other OECD countries, and the Gini coefficient has changed little between 2001 and 2019 [@australian2018rising], suggesting income inequality has remained steady over this time period. Our results do not conflict with this conclusion. Rather we are revealing that even a static income distribution may have dynamic effects on happiness over time. This highlights the issue that while traditional univariate measures of wealth and income inequality may be relatively stable and exhibit little change, their impact on wellbeing and health can still vary. As such, these results may well have relevance to other developed nations in North America and Europe which also enjoy low levels of income inequality, but have stagnating incomes, and declining happiness levels [@sachs2019world]. Establishing the links between income, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical aim as government begin to shift focus away from traditional measures of economic prosperity.

>>>>

A fundamental question for society is how much happiness does a dollar buy? The accepted view among economists and psychologists is that money and happiness increase together up to a point, after which there is little further gain from increasing income. While the location of this change point reportedly ranges between USD\$60K to \$95K, there has been no investigation as to whether this has increased or decreased over time. We tested the temporal relationship between income and affective wellbeing (happiness), and income and cognitive wellbeing (life satisfaction), using household economic data from Australia between 2001-2019. We discovered the change point between happiness and income has increased over those 19 years faster than inflation (i.e., cost of living). This suggests that inequalities in income may be driving increasing inequities in happiness between the rich and the poor, with implications for health and recent government policy-goals to monitor and improve wellbeing.

@@

Australia has low levels of income disparity relative to many other OECD countries, and the Gini coefficient has changed little between 2001 and 2019 [@australian2018rising], suggesting income inequality has remained steady over this time period. Our results do not conflict with this conclusion. Rather we are revealing that even a static income distribution may have dynamic effects on happiness over time. This highlights the issue that while traditional univariate measures of income and income inequality may be relatively stable and exhibit little change, their impact on wellbeing and health can still vary. As such, these results may well have relevance to other developed nations in North America and Europe which also enjoy low levels of income inequality, but have stagnating incomes, and declining happiness levels [@sachs2019world]. Establishing the links between income, wellbeing and health, and how inequalities in one drives inequities in the other, will be a critical aim as government begin to shift focus away from traditional measures of economic prosperity.
